A Logic-Based Framework for Digital Signal Processing

### 2008-05-27 trunk/paper/paper.tex

\documentclass{llncs}
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{url}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

@@

\title{A logic-based framework for Digital Signal Processing}

\author{...}

@@

\begin{abstract}
In this paper we describe a logic-based framework for Digital Signal Processing. This framework is implemented on SWI-Prolog and is formed of interfaces with different DSP libraries and APIs extending this logic language ``to speak'' DSP. This approach presents the advantage, in respect of other integrated APIs, of being application implementation independent, so the extended logic language can be loaded an embedded in any application or language interfacing SWI-Prolog. The new DSP capabilities are public as predicates grouped in modules which offer different levels of abstraction. Low-level modules offer predicates to communicate directly with the API. Task-oriented modules define some default routines which specific goals and procedures. Finally we can relate input/output through ``black boxes'' representing a semantic relationship between them. Modules can communicate by using commong semantic data types declared as functors which determine the possible workflows. Such a framework allow us to build complex DSP queries as procedures as well as simplify the extraction of audio processing knowledge with common semantics. Finally, Concurrent Transaction Logic allows us to wrap the procedural meaning of predicate rules into a declarative definition of transactions controlling the execution and updating of the processes.
\emph{abstract} environment.

\end{abstract}

\section{Introduction}

Digital signal processing is a broad paradigm and object of an important research. Within this community there are many particular investigators discovering and developing new algorithms for signal processing and analysis. Many projects have been directed to develop formal APIs providing a framework for algorithms implementation and publication. However these APIs are normally designed as a close environment which requires a specific developer knowledge or as a ``plugin'' library which requires the implementation of a specific host.

This situation leads to many isolated applications, environments and sources of data which work independently from each other. Any effort to combine them to obtain integrated results for analysis, classification and comparison requires tedious studies or ``glue code'' implementations.

In this paper we address this problem by proposing and developing SWI-DSP, a logical framework for \textit{Audio Processing} that extends the SWI-Prolog vocabulary. We rely on a logic programming language with audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end which is generally more understandable than a closed implementation in any Procedural or Object Oriented language. In addition to this, we are keen on using logic resources from SWI-Prolog to obtain finally a more capable DSP engine.

In \S 2 we summarize and evaluate the alternative APIs for audio analysis. Some of these APIs are hosted in SWI-DSP through the foreign interface providing an open system for developers as well. In \S 3 we start viewing the system from the top and how it defines an interesting workflow which actually solves the most part of the problems of the DSP paradigm. This workflow is determined by semantic types, a binary data standard and procedural declarations of riles. \S4 goes into detail in each task-oriented module describing the communication layers with the APIs through control predicates which require an understanding of the API.

Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining semantic types and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction. However, this front-end still requires specific SWI-Prolog knowledge to create programs suitable for the concrete needs and it does not offer any efficient tabling mechanism. \S 5 explains how Concurrent Transaction Logic is the solution to provide a better logic front-end combining procedural and declarative meaning of transactions which also comprise data updating. In \S 6 we present related works as alternatives to manage SWI-DSP featuring another user interface than SWI-Prolog. We conclude in \S 7 by assessing the value and scope of this work and introducing future advances.

\section{Audio Analysis tasks and solutions}

Audio analysis  APIs offer formal implementations for development and use of different algorithms and techniques. These algorithms are focused on different tasks that can be classified in these 3 groups:

\begin{itemize}
 \item Processing
 \item Feature Extraction
 \item Classification
\end{itemize}

\subsection{Audio Processing}

@@

One famous API is the Linux Audio Developer's Simple Plugin API (LADSPA) which has been succesfully used in many systems and applications and is a widely spread implementation for algorithms development.

\subsection{Feature Extraction}

Feature extraction is an important paradigm in signal research as it focuses on extract qualitative characteristics of the signal in a different domain than the input. Features are classified in different levels of abstraction from frequency characteristics to high-level audio ones.

The Vamp Plugins API offers another standard and utilities set to develop feature extraction plugins that can be hosted in any application able to control the ``plugin lifecycle''.

\subsection{Classification}

From different features we are able to classify different signals.

\subsection{Integrative solutions}

With the same motivation of this work there are some integrative APIs which combines different tasks to perform high-level operations in a closed framework. In the particular domain of Music jMIR and Marsyas present a complex framework for audio analysis and classification.

\section{SWI-DSP workflow}

@@

\subsection{Semantic datatypes}

@@

\subsection{Large Binary Objects}

@@

\subsection{Workflow}

@@

KeyProfile = ['Feature'(key, 'Timestamp'(0, 23.45), 'Amajor`')|...].

@@

\section{SWI-DSP modules}

Each task module is composed by lower level modules which offer communication with an external API or library in charge of the computations. By hosting APIs we can extend easily SWI-DSP vocabulary as we do not need to assemble each algorithm.

\subsection{swiaudio}

@@

\subsection{feature extraction}

This module defines a standard task-oriented set of predicates for feature extraction. As we have seen in \S 2 the module is characterized for a clear input/output relationship. So we can declare a simple predicate to extract the tempo of an audio signal: tempo(+Input, -Tempo) being Input an instantation of the Signal datatype and Tempo a list of instantations of the Feature datatype. This ``magic'' predicate is triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo}.

@@

The sub-module \textbf{swivamp} is a communication layer based on the SWI-Prolog foreign interface which allow us to query the plugin information, set up the plugin context for the transform, execute the plugin given a Signal term and retrieve the output features for the plugin according to the lifecycle specification. Plugins are wrapped in a similar fashion than the audio data so we can use a plugin instance representation which allow us to have concurrent plugins working at the same time. The previous tempo/3 predicate is expressed as this set of rules:

@@

\subsection{audio processing}

@@

\subsection{swiweka}

\section{Towards a computation reasoner}

@@

\subsection{DSP scenarios}

@@

\subsection{Computation reasoning}

@@

\section{SWI-DSP in real applications}

@@

\subsection{Henry}

@@

 \section{Conclusion and Future work}

>>>>

\documentclass[runningheads]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\usepackage{url}
\urldef{\mails}\path|{david.pastor,yves.raimond, mark.sandler}@elec.qmul.ac.uk|
%\newcommand{\keywords}[1]{\par\addvspace\baselineskip
%\noindent\keywordname\enspace\ignorespaces#1}

@@

\mainmatter

\title{A Logic-Based Framework for Digital Signal Processing}

\titlerunning{A Logic-Based Framework for Digital Signal Processing}

\author{David Pastor\and Yves Raimond\and Mark Sandler}

\institute{Centre for Digital Music, Queen Mary, University of London,\\
\mails}

@@

\begin{abstract}
In this paper we describe a logic-based framework for Digital Signal Processing. This framework is implemented on SWI-Prolog and is formed of interfaces with different DSP libraries and APIs extending this logic language ``to speak'' DSP. This approach presents the advantage, in respect of other integrated APIs, of being application implementation independent, so the extended logic language can be loaded an embedded in any application or language interfacing SWI-Prolog. The new DSP capabilities are public as predicates grouped in modules which offer different levels of abstraction. Low-level modules offer predicates to communicate directly with the API. Task-oriented modules define some default routines which specific goals and procedures. Finally we can relate input/output through ``black boxes'' representing a semantic relationship between them. Modules can communicate by using commong semantic data types declared as functors which determine the possible workflows. Such a framework allow us to build complex DSP queries as procedures as well as simplify the extraction of audio processing knowledge with common semantics. Finally, Concurrent Transaction Logic allows us to wrap the procedural meaning of predicate rules into a declarative definition of transactions controlling the execution and updating of the processes.
\keywords{DSP predicates, computation query, computation reasoner, DSP scenarios, RDF}
\end{abstract}

\section{Introduction}\label{sec:intro}

Digital signal processing is a broad field and an object of an important research. Withing this research community, particular investigators are discovering and developing new algorithms for signal processing and analysis. Many projects aim at developing formal APIs providing a framework for algorithms implementation and publication. However, these APIs are normally designed as a closed environment which hardly interacts with other related APIs and applications. Although several APIs are wrapped and exported as ``plugin'' libraries which can be embedded in different applications, they still require the implementation of specific and non-interactive hosts.

This situation leads to many isolated applications, environments and sources of data working independently from each other. Any effort to combine them to obtain integrated results for analysis, classification and comparison requires tedious studies or ``glue code'' implementations.

In this paper we address this problematic situation by proposing and developing SWI-DSP, a logical framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog vocabulary. We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end which is generally more understandable than a closed implementation in any Procedural or Object Oriented language. In addition to this, we are keen on using logic resources to finally achieve a more capable DSP engine.

In \S 2 we identify the most common audio analysis tasks and summarize several relevant \textit{open source} APIs aimed at performing such tasks. We evaluate the advantages and limitations of these APIs in order to find the requirements for SWI-DSP. In \S 3 we start viewing the system from the top and how it defines an interesting workflow that characterizes input/output relationships between task-oriented modules. This workflow is determined by semantic types, a binary data standard and procedural declarations of DSP rules. \S 4 goes into the implementation of each task-oriented module which are indeed communication layers with dedicated ``plugin'' libraries that are in charge of the actual computation and offer an open development framework.

Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining semantic types and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction. However, this front-end does not present an efficient and manageable framework as it requires specific SWI-Prolog knowledge to create programs suitable for the concrete needs and it does not offer any efficient tabling mechanism. We tackle this problem in \S 5 by defining DSP ``scenarios'' which offer a semantic view of the system to the user without losing computational flexibility.

We also describe as a Standard, how Concurrent Transaction Logic is the solution to provide an efficient engine allowing a formal data updating mechanism and computation reasoning capabilities. Finally, \S 6 presents related works that use SWI-DSP to create an environment for multimedia data management and visualisation. We conclude in \S 7 by assessing the value and scope of this work and introducing future advances.

\section{Audio Analysis tasks and solutions}\label{sec:tasks}

Audio analysis and synthesis APIs offer formal implementations for development and use of different algorithms and techniques. These algorithms are focused on different tasks that can be classified in these 3 groups:

\begin{itemize}
 \item Processing and Effects
 \item Feature Extraction
 \item Classification
\end{itemize}

\subsection{Audio Processing}\label{subsec:effects}

@@

One important dedicated API is the Linux Audio Developer's Simple Plugin API (LADSPA) which has been succesfully used in many systems and applications and is a widely spread implementation for algorithms development. We host LADSPA in SWI-DSP but there are other interesting libraries like DSSI plugins for effects and LV2 (new generation of LADSPA) which are targets for future extensions. (this is too simple i think)

\subsection{Feature Extraction}\label{subsec:feature}

Feature extraction is an important paradigm in signal research as it focuses on extract qualitative characteristics of the signal in a different domain than the input. Features are classified in different levels of abstraction from frequency and time domain characteristics to high-level audio ones.

The Vamp Plugins API offers a standard and utilities to develop feature extraction plugins that can be hosted in any application able to control the ``plugin lifecycle''. We host Vamp plugins as basis for feature extraction in SWI-DSP. In general plugin libraries are characterized for being easy for development but application-dependent. By embedding plugins in SWI-DSP we offer a flexible development framework and make an interactive interface to deal with plugins in several ways.

\subsection{Integrative solutions}\label{subsec:inte}

With the same motivation of this work there are some integrative APIs which combines different tasks to perform low high-level operations in a closed implementation. CLAM, jMIR and Marsyas present a complex framework collecting tools to perform different DSP tasks. The most interesting feature of these frameworks is the possibility to create DSP networks combining specific algorithms and modules thanks to a workflow design.

CLAM also offers a metamodel of DSP processes wich provides semantics to design the networks properly presenting a higher and more manageable level of abstraction. However, these APIs still present two important limitations that we pretend to solve:

\begin{itemize}
 \item There do not offer a common format to exchange data or an easy way to interact with other applications.
 \item They offer a procedural system unable to reach higher-level knowledge by itself.
\end{itemize}

\subsection{Classification}\label{subsec:classif}

Classification is one the most interesting and complex paradigms in any specific domain exploiting the ``intelligence'' of computational machines. The WEKA API collects different machine-learning algorithms and provides a formal interface (in Java) for development and interaction with other systems.

However, this generic API does not provide semantics to create a dataset which is the most intriguing and complex issue for classifying data properly. Marsyas and jMÍR integrate WEKA orienting it for music classification purporses by offering a code interface with other modules (Marsyas) or by extending the WEKA file format ARFF (jMIR). Both solutions still seem to lack of effective semantics to perform classification of audio data.

\section{SWI-DSP workflow}\label{sec:dspworkflow}

@@

\subsection{Semantic datatypes}\label{subsec:datatypes}

@@

\subsection{Large Binary Objects}\label{subsec:blobs}

@@

\subsection{Workflow}\label{subsec:worflow}

@@

KeyProfile = ['Feature'(key, 'Timestamp'(0, 23.45), 'Amajor')|...].

@@

\section{SWI-DSP modules}\label{sec:modules}

Each task module is composed by lower level modules which offer communication with an external API or library in charge of the computations. By hosting ``plugin libraries'' we can extend easily SWI-DSP vocabulary as we do not need to assemble each algorithm.

\subsection{swiaudio}\label{subsec:swiaudio}

@@

\subsection{feature extraction}\label{subsec:swivamp}

This module defines a standard task-oriented set of predicates for feature extraction. As we have seen in \S 2 the module is characterized for a clear input/output relationship. So we can declare a simple predicate to extract the tempo of an audio signal: tempo(+Input, -Tempo) being Input an instantation of the Signal datatype and Tempo a list of instantations of the Feature datatype. This ``magic'' predicate is triggering a specific computation tool, so we could select the specific tool by extending the predicate: \textit{tempo(+Input, +Tool, -Tempo)}.

@@

The sub-module \textbf{swivamp} is a communication layer based on the SWI-Prolog foreign interface which allow us to query the plugin information, set up the plugin context for the transform, execute the plugin given a Signal term and retrieve the output features for the plugin according to the lifecycle specification. Plugins are wrapped in a similar fashion than the audio data so we can use a plugin instance representation which allow us to have concurrent plugins working at the same time. The previous tempo/3 predicate is declared as this set of rules:

@@

\subsection{audio processing}\label{subsec:swilasdpa}

@@

\subsection{swiweka}\label{subsec:swiweka}

\section{Towards a computation reasoner}\label{sec:reasoner}

@@

\subsection{DSP scenarios}\label{subsec:scenarios}

@@

\subsection{Computation reasoning}\label{subsec:ctl}

@@

\section{SWI-DSP in real applications}\label{sec:realapp}

@@

\subsection{Henry}\label{subsec:henry}

@@

 \section{Conclusion and Future work}\label{sec:conclusion}
